---
title: "Species distribution modeling"
author: "Tim Richter"
date: "25 8 2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Species distribution modeling of Phylloscopus trochilus (willow warbler) in Europe under current and future climate conditions.

```{r}
rm(list=ls())

library(raster)
```

# ABUNDANCE DATA PRE-PROCESSING#

```{r}
abundancepath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Abundance_data/"

abund <- read.csv2(paste0(abundancepath, "willow_warbler_2020_Europe_latlong.csv"), header=T, sep=";")

abund$MONTH <- substr((abund$OBSERVATION.DATE), 4, 5)

ww <- abund[abund$MONTH == "06"| abund$MONTH == "05",]
```
## Checking for errors/NAs
```{r}
lonzero <- subset(ww, ww$LONGITUDE==0)
latzero <- subset(ww, ww$LATITUDE==0)
# --> no errors(NAs)
```
## Checking for exact duplicates
```{r}
dups <- duplicated(ww) # logical string e.g. [FALSE FALSE FALSE TRUE FALSE TRUE...]
wwdups <- ww[dups, ]
# --> no duplicates
```
## Checking for same coordinate data
```{r}
dups2 <- duplicated(wwdups[, c("LATITUDE", "LONGITUDE")])
wwnd <- ww[!dups2,] # Subsample with no coordinate duplicates
```
## Saving as .csv
```{r}
setwd(abundancepath)
write.csv2(ww, file="willow_warbler_2020_Europe_prep_and_clean.csv", row.names=TRUE)
```

# Reprojection #
```{r}
rm(list=ls())

library(raster)
library(dismo)
library(sf)
library(sp)
library(mapview)

predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"
abundancepath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Abundance_data/"
```
## PROJECT OCCURRENCE COORDINATES
```{r}
ww <- read.csv2(paste0(abundancepath, "willow_warbler_2020_Europe_prep_and_clean.csv"), header=T, sep=";")

currentclim <- raster::stack(paste0(predfinpath, "currentclim.grd"))

coordinates(ww)=~LONGITUDE+LATITUDE
proj4string(ww) <- crs("+init=epsg:4326")
wwETRS <- spTransform(ww, crs(currentclim))
```
## ADD PROJECTED LONGITUDE LATITUDE COLUMNS
```{r}
wwlonlat <- data.frame(wwETRS)
```
## CREATE SUBSET WITH ONLY LONGITUDE AND LATITUDE COLUMNS
```{r}
wwsdm <- subset(wwlonlat, select=c("LONGITUDE", "LATITUDE"))
```
## SAVE DATAFRAME
```{r}
setwd(abundancepath)

saveRDS(wwsdm, "wwfin.RDS")
```

# PREDICTORS PRE-PROCESSING #
```{r}
library(mapview)
library(stringr)
```
## Worldclim monthly averages from 2010-2018
```{r}
preclist <- list.files("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_prec_2010-2018/", pattern="*.tif$")
tminlist <- list.files("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_tmin_2010-2018/", pattern="*.tif$")
tmaxlist <- list.files("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_tmax_2010-2018/", pattern="*.tif$")

prec <- raster::stack(paste0("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_prec_2010-2018/", preclist))
tmin <- raster::stack(paste0("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_tmin_2010-2018/", tminlist))
tmax <- raster::stack(paste0("E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/Current_climate/wc2.1_2.5m_tmax_2010-2018/", tmaxlist))
```
## Calculating means for every month

## PRECIPITATION

## Creating subsets for each month 
```{r}
prec01 <- raster::subset(prec, grep(".01", names(prec), fixed=TRUE, value=T))
prec02 <- raster::subset(prec, grep(".02", names(prec), fixed=TRUE, value=T))
prec03 <- raster::subset(prec, grep(".03", names(prec), fixed=TRUE, value=T))
prec04 <- raster::subset(prec, grep(".04", names(prec), fixed=TRUE, value=T))
prec05 <- raster::subset(prec, grep(".05", names(prec), fixed=TRUE, value=T))
prec06 <- raster::subset(prec, grep(".06", names(prec), fixed=TRUE, value=T))
prec07 <- raster::subset(prec, grep(".07", names(prec), fixed=TRUE, value=T))
prec08 <- raster::subset(prec, grep(".08", names(prec), fixed=TRUE, value=T))
prec09 <- raster::subset(prec, grep(".09", names(prec), fixed=TRUE, value=T))
prec10 <- raster::subset(prec, grep(".10", names(prec), fixed=TRUE, value=T))
prec11 <- raster::subset(prec, grep(".11", names(prec), fixed=TRUE, value=T))
prec12 <- raster::subset(prec, grep(".12", names(prec), fixed=TRUE, value=T))
prec12 <- dropLayer(prec12, 10) # 2019 Data should not be considered

prec01mean <- calc(prec01, mean) # calculating mean -> 1 raster layer
prec02mean <- calc(prec02, mean)
prec03mean <- calc(prec03, mean)
prec04mean <- calc(prec04, mean)
prec05mean <- calc(prec05, mean)
prec06mean <- calc(prec06, mean)
prec07mean <- calc(prec07, mean)
prec08mean <- calc(prec08, mean)
prec09mean <- calc(prec09, mean)
prec10mean <- calc(prec10, mean)
prec11mean <- calc(prec11, mean)
prec12mean <- calc(prec12, mean)

precmean <- stack(prec01mean, prec02mean, prec03mean, prec04mean, prec05mean, prec06mean, prec07mean, prec08mean, prec09mean, prec10mean, prec11mean, prec12mean)



predpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors/"
setwd(predpath)
writeRaster(precmean, "precmean.tif",format="GTiff", overwrite=TRUE)
```
## MINIMUM TEMPERATURE

## Creating subsets for each month 
```{r}
tmin01 <- raster::subset(tmin, grep(".01", names(tmin), fixed=TRUE, value=T))
tmin02 <- raster::subset(tmin, grep(".02", names(tmin), fixed=TRUE, value=T))
tmin03 <- raster::subset(tmin, grep(".03", names(tmin), fixed=TRUE, value=T))
tmin04 <- raster::subset(tmin, grep(".04", names(tmin), fixed=TRUE, value=T))
tmin05 <- raster::subset(tmin, grep(".05", names(tmin), fixed=TRUE, value=T))
tmin06 <- raster::subset(tmin, grep(".06", names(tmin), fixed=TRUE, value=T))
tmin07 <- raster::subset(tmin, grep(".07", names(tmin), fixed=TRUE, value=T))
tmin08 <- raster::subset(tmin, grep(".08", names(tmin), fixed=TRUE, value=T))
tmin09 <- raster::subset(tmin, grep(".09", names(tmin), fixed=TRUE, value=T))
tmin10 <- raster::subset(tmin, grep(".10", names(tmin), fixed=TRUE, value=T))
tmin11 <- raster::subset(tmin, grep(".11", names(tmin), fixed=TRUE, value=T))
tmin12 <- raster::subset(tmin, grep(".12", names(tmin), fixed=TRUE, value=T))
tmin12 <- dropLayer(tmin12, 10) # 2019 Data should not be considered

tmin01mean <- calc(tmin01, mean) # calculating mean -> 1 raster layer
tmin02mean <- calc(tmin02, mean)
tmin03mean <- calc(tmin03, mean)
tmin04mean <- calc(tmin04, mean)
tmin05mean <- calc(tmin05, mean)
tmin06mean <- calc(tmin06, mean)
tmin07mean <- calc(tmin07, mean)
tmin08mean <- calc(tmin08, mean)
tmin09mean <- calc(tmin09, mean)
tmin10mean <- calc(tmin10, mean)
tmin11mean <- calc(tmin11, mean)
tmin12mean <- calc(tmin12, mean)

tminmean <- stack(tmin01mean, tmin02mean, tmin03mean, tmin04mean, tmin05mean, tmin06mean, tmin07mean, tmin08mean, tmin09mean, tmin10mean, tmin11mean, tmin12mean)



predpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors/"
setwd(predpath)
writeRaster(tminmean, "tminmean.tif",format="GTiff", overwrite=TRUE)
```

## MAXIMUM TEMPERATURE

## Creating subsets for each month 
```{r}
tmax01 <- raster::subset(tmax, grep(".01", names(tmax), fixed=TRUE, value=T))
tmax02 <- raster::subset(tmax, grep(".02", names(tmax), fixed=TRUE, value=T))
tmax03 <- raster::subset(tmax, grep(".03", names(tmax), fixed=TRUE, value=T))
tmax04 <- raster::subset(tmax, grep(".04", names(tmax), fixed=TRUE, value=T))
tmax05 <- raster::subset(tmax, grep(".05", names(tmax), fixed=TRUE, value=T))
tmax06 <- raster::subset(tmax, grep(".06", names(tmax), fixed=TRUE, value=T))
tmax07 <- raster::subset(tmax, grep(".07", names(tmax), fixed=TRUE, value=T))
tmax08 <- raster::subset(tmax, grep(".08", names(tmax), fixed=TRUE, value=T))
tmax09 <- raster::subset(tmax, grep(".09", names(tmax), fixed=TRUE, value=T))
tmax10 <- raster::subset(tmax, grep(".10", names(tmax), fixed=TRUE, value=T))
tmax11 <- raster::subset(tmax, grep(".11", names(tmax), fixed=TRUE, value=T))
tmax12 <- raster::subset(tmax, grep(".12", names(tmax), fixed=TRUE, value=T))
tmax12 <- dropLayer(tmax12, 10) # 2019 Data should not be considered

tmax01mean <- calc(tmax01, mean) # calculating mean -> 1 raster layer
tmax02mean <- calc(tmax02, mean)
tmax03mean <- calc(tmax03, mean)
tmax04mean <- calc(tmax04, mean)
tmax05mean <- calc(tmax05, mean)
tmax06mean <- calc(tmax06, mean)
tmax07mean <- calc(tmax07, mean)
tmax08mean <- calc(tmax08, mean)
tmax09mean <- calc(tmax09, mean)
tmax10mean <- calc(tmax10, mean)
tmax11mean <- calc(tmax11, mean)
tmax12mean <- calc(tmax12, mean)

tmaxmean <- stack(tmax01mean, tmax02mean, tmax03mean, tmax04mean, tmax05mean, tmax06mean, tmax07mean, tmax08mean, tmax09mean, tmax10mean, tmax11mean, tmax12mean)

predpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors/"
setwd(predpath)
writeRaster(tmaxmean, "tmaxmean.tif",format="GTiff", overwrite=TRUE)
```

# PREDICTORS STANDARDIZATION AND MERGING #
```{r}
rm(list=ls())

library(raster)
library(dismo)
library(sf)
library(mapview)
```
## Loading predictors
```{r}
gadmpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/GADM_Europa/"
predpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors/"
corpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Corine Land Cover/DATA/"
predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"
elepath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/WorldClim/wc2.1_2.5m_elev/"
```
## In ETRS89 umprojizieren
## WORLDCLIM

## current
```{r}
precmean <- stack(paste0(predpath, "precmean.grd"))
tmaxmean <- stack(paste0(predpath, "tmaxmean.grd"))
tminmean <- stack(paste0(predpath, "tminmean.grd"))


precmeanETRS <- projectRaster(precmean, crs="+init=epsg:3035")
tmaxmeanETRS <- projectRaster(tmaxmean, crs="+init=epsg:3035")
tminmeanETRS <- projectRaster(tminmean, crs="+init=epsg:3035")

setwd(predpath)
writeRaster(precmeanETRS, "precmeanETRS.grd", overwrite=TRUE)
writeRaster(tmaxmeanETRS, "tmaxmeanETRS.grd", overwrite=TRUE)
writeRaster(tminmeanETRS, "tminmeanETRS.grd", overwrite=TRUE)

climmeanETRS <- raster::stack(precmeanETRS, tmaxmeanETRS, tminmeanETRS)

setwd(predpath)
writeRaster(climmeanETRS, "climmeanETRS.grd", overwrite=TRUE)
```
## CORINE LAND COVER
```{r}
cor <- raster(paste0(corpath, "U2018_CLC2018_V2020_20u1.tif"))
```
## Creating Rastertemplate
```{r}
precmeanETRS <- stack(paste0(predpath, "precmeanETRS.grd"))
tmaxmeanETRS <- stack(paste0(predpath, "tmaxmeanETRS.grd"))
tminmeanETRS <- stack(paste0(predpath, "tminmeanETRS.grd"))

climmeanETRS <- raster::stack(paste0(predpath, "climmeanETRS.grd"))

rastemplate <- raster(ext = extent(cor), crs = crs(cor), resolution = c(10000, 10000))
```
## Cropping
```{r}
precmeancrop <- crop(precmeanETRS, cor)
tmaxmeancrop <- crop(tmaxmeanETRS, cor)
tminmeancrop <- crop(tminmeanETRS, cor)
```
## Resampling
```{r}
precmeanfin <- resample(precmeancrop, rastemplate)
tmaxmeanfin <- resample(tmaxmeancrop, rastemplate)
tminmeanfin <- resample(tminmeancrop, rastemplate)

climmeanfin <- raster::stack(precmeanfin, tmaxmeanfin, tminmeanfin)

setwd(predfinpath)
writeRaster(climmeanfin, "climmeanfin.grd", overwrite=TRUE)
```
## future
```{r}
library(raster)
library(dismo)
library(sf)
library(mapview)
```
## Loading predictors
```{r}
gadmpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/GADM_Europa/"
predpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors/"
corpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Corine Land Cover/DATA/"
predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"

Prec245 <- stack(paste0(predpath, "Prec245.grd"))
Prec370 <- stack(paste0(predpath, "Prec370.grd"))
Tmax245 <- stack(paste0(predpath, "Tmax245.grd"))
Tmax370 <- stack(paste0(predpath, "Tmax370.grd"))
Tmin245 <- stack(paste0(predpath, "Tmin245.grd"))
Tmin370 <- stack(paste0(predpath, "Tmin370.grd"))
```
## rastertemplate
```{r}
climmean <- stack(paste0(predfinpath, "climmeanfin.grd"))

precmean <- stack(paste0(predpath, "precmean.grd"))

rastemplate <- projectRaster(climmean, crs=crs(precmean))
```
## crop
```{r}
Prec245crop <- crop(Prec245, rastemplate)
Prec370crop <- crop(Prec370, rastemplate)
Tmax245crop <- crop(Tmax245, rastemplate)
Tmax370crop <- crop(Tmax370, rastemplate)
Tmin245crop <- crop(Tmin245, rastemplate)
Tmin370crop <- crop(Tmin370, rastemplate)
```
## project
```{r}
Prec245ETRS <- projectRaster(Prec245crop, crs="+init=epsg:3035")
Prec370ETRS <- projectRaster(Prec370crop, crs="+init=epsg:3035")
Tmax245ETRS <- projectRaster(Tmax245crop, crs="+init=epsg:3035")
Tmax370ETRS <- projectRaster(Tmax370crop, crs="+init=epsg:3035")
Tmin245ETRS <- projectRaster(Tmin245crop, crs="+init=epsg:3035")
Tmin370ETRS <- projectRaster(Tmin370crop, crs="+init=epsg:3035")

setwd(predpath)
writeRaster(Prec245ETRS, "Prec245ETRS.grd", overwrite=TRUE)
writeRaster(Prec370ETRS, "Prec370ETRS.grd", overwrite=TRUE)
writeRaster(Tmax245ETRS, "Tmax245ETRS.grd", overwrite=TRUE)
writeRaster(Tmax370ETRS, "Tmax370ETRS.grd", overwrite=TRUE)
writeRaster(Tmin245ETRS, "Tmin245ETRS.grd", overwrite=TRUE)
writeRaster(Tmin370ETRS, "Tmin370ETRS.grd", overwrite=TRUE)

futureclimETRS <- raster::stack(Prec245ETRS, Prec370ETRS, Tmax245ETRS, Tmax370ETRS, Tmin245ETRS, Tmin370ETRS)

writeRaster(futureclimETRS, "futureclimETRS.grd", overwrite=TRUE)
```
## resample
```{r}
Prec245fin <- resample(Prec245ETRS, climmean)
Prec370fin <- resample(Prec370ETRS, climmean)
Tmax245fin <- resample(Tmax245ETRS, climmean)
Tmax370fin <- resample(Tmax370ETRS, climmean)
Tmin245fin <- resample(Tmin245ETRS, climmean)
Tmin370fin <- resample(Tmin370ETRS, climmean)

futureclim245 <- raster::stack(Prec245fin, Tmax245fin, Tmin245fin)
futureclim370 <- raster::stack(Prec370fin, Tmax370fin, Tmin370fin)

setwd(predfinpath)
writeRaster(futureclim245, "futureclim245.grd", overwrite=TRUE)
writeRaster(futureclim370, "futureclim370.grd", overwrite=TRUE)
```
## RESAMPLE CORINE LAND COVER
```{r}
cor10km <- aggregate(cor, fact=100, fun=modal) # this is necessary because its a categorical variable

corfin <- resample(cor10km, climmean, method="ngb")

setwd(predfinpath)
writeRaster(corfin, "corfin.grd", overwrite=TRUE)

## categorical resampling:
# http://citeseerx.ist.psu.edu/viewdoc/download;jsessionid=E54988E2300738F35AFC8DEE2E261B32?doi=10.1.1.139.9293&rep=rep1&type=pdf
```
## Resample Elevation
```{r}
ele <- raster(paste0(elepath, "wc2.1_2.5m_elev.tif"))

elecrop <- crop(ele, rastemplate)

eleETRS <- projectRaster(elecrop, crs="+init=epsg:3035")

elefin <- resample(eleETRS, climmean)

setwd(predfinpath)
writeRaster(elefin, "elevation.grd", overwrite=TRUE)
```
## Merging all the predictors
```{r}
predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"

futureclim245 <- raster::stack(paste0(predfinpath, "futureclim245.grd"))
futureclim370 <- raster::stack(paste0(predfinpath, "futureclim370.grd"))
corfin <- raster(paste0(predfinpath, "corfin.grd"))
elefin <- raster(paste0(predfinpath, "elevation.grd"))

currentclimstack <- raster::stack(climmean, corfin, elefin)
futureclim245stack <- raster::stack(futureclim245, corfin, elefin)
futureclim370stack <- raster::stack(futureclim370, corfin, elefin)

setwd(predfinpath)
writeRaster(currentclimstack, "currentclimstack.grd", overwrite=TRUE)
writeRaster(futureclim245stack, "futureclim245stack.grd", overwrite=TRUE)
writeRaster(futureclim370stack, "futureclim370stack.grd", overwrite=TRUE)
```
## Mask predictors to GADM
```{r}
predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"
gadmpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/GADM_Europa/"

currentclimstack <- raster::stack(paste0(predfinpath, "currentclimstack.grd"))
futureclim245stack <- raster::stack(paste0(predfinpath, "futureclim245stack.grd"))
futureclim370stack <- raster::stack(paste0(predfinpath, "futureclim370stack.grd"))

eugadm<- st_read(paste0(gadmpath, "Europe.shp")) # Quelle siehe "Readme", https://tapiquen-sig.jimdofree.com/english-version/free-downloads/europe/

eugadmgrs <- st_transform(eugadm, crs(currentclimstack))

currentclimstackeu <- mask(currentclimstack, eugadmgrs)
futureclim245stackeu <- mask(futureclim245stack, eugadmgrs)
futureclim370stackeu <- mask(futureclim370stack, eugadmgrs)

setwd(predfinpath)
writeRaster(currentclimstackeu, "currentclimstack.grd", overwrite=TRUE)
writeRaster(futureclim245stackeu, "futureclim245stack.grd", overwrite=TRUE)
writeRaster(futureclim370stackeu, "futureclim370stack.grd", overwrite=TRUE)

```
# MODEL TRAINING, PREDICTIONS AND NORTHERLY SHIFT #
```{r}
rm(list=ls())

library(raster)
library(dismo)
library(rJava)
library(sf)
library(sp)
library(mapview)

predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"
abundancepath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Abundance_data/"
modelpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Models/"
```
## LOAD DATA
```{r}
ww <- readRDS(paste0(abundancepath, "wwfin.RDS"))

currentclimstack <- raster::stack(paste0(predfinpath, "currentclimstack.grd"))
futureclim245stack <- raster::stack(paste0(predfinpath, "futureclim245stack.grd"))
futureclim370stack <- raster::stack(paste0(predfinpath, "futureclim370stack.grd"))
```
## MODEL TRAINING
## Create background points
```{r}
# install.packages("blockCV", dependencies=TRUE)
# remotes::install_github("rvalavi/blockCV", dependencies=TRUE)
library(maxnet)
library(precrec)
library(blockCV)

mask <- currentclimstack$Prec.1
# select 10000 random points
# set seed to assure that the examples will always
# have the same random sample
set.seed(1963)

bg <- randomPoints(mask, 10000, p = ww)

bgframe <- data.frame(bg)

gadmpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/GADM_Europa/"

eugadm<- st_read(paste0(gadmpath, "Europe.shp")) # Quelle siehe "Readme", https://tapiquen-sig.jimdofree.com/english-version/free-downloads/europe/
eugadmgrs <- st_transform(eugadm, crs(currentclimstack))

plot(eugadmgrs$geometry, main = "Presence data of phylloscopus trochilus")
points(ww)

plot(eugadmgrs$geometry, main = "Randomly sampled background points")
points(bgframe)

sdm <- maxent(currentclimstack, p = ww, a = bgframe, factors=37, removeDuplicates=TRUE)

setwd(modelpath)

saveRDS(sdm, "maxent.RDS")
```
## PREDICTION
```{r}
sdm <- readRDS(paste0(modelpath, "maxent.RDS"))
sdmcurrent <- predict(sdm, currentclimstack)

plot(sdmcurrent, main = "Willow warbler predicted current distribution", xlab = "Easting", ylab="Northing", sub = "The color scheme represents the distribution probability")

sdm245 <- predict(sdm, futureclim245stack)
sdm370 <- predict(sdm, futureclim370stack)

plot(sdm245, main = "Willow warbler predicted distribution for the years 2041-2060 (ssp245)", xlab="Easting", ylab="Northing", sub = "The color scheme represents the distribution probability")
plot(sdm370, main = "Willow warbler predicted distribution for the years 2041-2060 (ssp370)", xlab="Easting", ylab="Northing", sub = "The color scheme represents the distribution probability")

par(mfrow=c(1,2))
```
## Shift in willow warbler distribution
```{r}
library(rgdal)
```
## Current climate
```{r}
par(mfrow=c(2,1))

abundprobcurrent <- as.vector(sdmcurrent$layer)
currentframe <- data.frame(abundprobcurrent)

currentframe$northing <- coordinates(sdmcurrent)[,2]

plot(currentframe$abundprobcurrent ~ currentframe$northing, type="h" ,main = "Willow warbler North-South distribution in the examined region - current climate", xlab="Northing", ylab="Distribution probability")
```
## Future climate
```{r}
abundprobfuture <- as.vector(sdm245$layer)
futureframe <- data.frame(abundprobfuture)

futureframe$northing <- coordinates(sdm245)[,2]

plot(futureframe$abundprobfuture ~ futureframe$northing, type="h",main = "Willow warbler North-South distribution in the examined region - future climate", xlab="Northing", ylab="Distribution probability")
```

### https://cran.r-project.org/web/packages/blockCV/vignettes/BlockCV_for_SDM.html

# BLOCK-CROSS VALIDATION #
```{r}
rm(list=ls())

library(raster)
library(dismo)
library(rJava)
library(sf)
library(sp)
library(mapview)
library(caret)
library(SDMtune)

predfinpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Predictors_finished/"
abundancepath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Abundance_data/"
modelpath <- "E:/Landschaftsökologie_Master/Module/M8a_Fernerkundung_und_raeumliche_Modellierung/Projekt_SDM/Data/Models/"
```
## LOAD DATA
```{r}
ww <- readRDS(paste0(abundancepath, "wwfin.RDS"))

currentclimstack <- raster::stack(paste0(predfinpath, "currentclimstack.grd"))
```
## Create background points
```{r}
library(maxnet)
library(precrec)
library(blockCV)
# install.packages("blockCV", dependencies=TRUE)
# remotes::install_github("rvalavi/blockCV", dependencies=TRUE)

mask <- currentclimstack$Prec.1
# select 10000 random points
# set seed to assure that the examples will always
# have the same random sample.
set.seed(1963)

bg <- randomPoints(mask, 10000, p = ww)

bgframe <- data.frame(bg)

bgframe$ABUNDANCE <- 0
ww$ABUNDANCE <- 1

names(bgframe) <- c("LONGITUDE", "LATITUDE", "ABUNDANCE")
wwframe <- rbind(ww, bgframe)

coordinates(wwframe)=~LONGITUDE+LATITUDE
proj4string(wwframe) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000
+y_0=3210000 +ellps=GRS80 +units=m +no_defs")

wwframecrop <- crop(wwframe, currentclimstack) # remove outliers
```
## create spatial blocks
```{r}
sb <- blockCV::spatialBlock(speciesData=wwframecrop,
                   species="ABUNDANCE",
                   rasterLayer = currentclimstack,
                   rows= 4,
                   cols= 10,
                   k = 5,
                   selection = "systematic")
```
## extract the raster values for the species points as a dataframe
```{r}
mydata <- raster::extract(currentclimstack, wwframecrop, na.rm=TRUE)
mydata <- as.data.frame(mydata)

wwlogical <- wwframecrop$ABUNDANCE

folds <- sb$foldID
```
## create an empty vector to store the AUC of each fold
```{r}
AUCs <- vector(mode = "numeric")
for(i in seq_len(38)){
  mydata[i][is.na(mydata[i])] <- 0
}
for(k in seq_len(5)){
  # extracting the training and testing indices
  # this way only works with foldID
  trainSet <- which(folds != k) # training set indices
  testSet <- which(folds == k) # testing set indices
  # fitting a maxent model using linear, quadratic and hinge features
  mx <- maxnet(p = wwlogical[trainSet], 
               data = mydata[trainSet, ], 
               maxnet.formula(p = wwlogical[trainSet], 
                              data = mydata[trainSet, ], 
                              classes = "default"))
  testTable <- wwframecrop[testSet, ] # a table for testing predictions and reference data
  testTable$pred <- predict(mx, mydata[testSet, ], type = "cloglog") # predict the test set
  # calculate area under the ROC curve
  # precrec_obj <- evalmod(scores = testTable$pred, labels = testTable$ABUNDANCE)
  # AUCs[k] <- auc(precrec_obj)[1,4] # extract AUC-ROC
}

# print(mean(AUCs))
```
